## Loading and preprocessing the data

```{r loadingdata, echo=TRUE}
require(plyr)
ActivityDF <- read.csv("~/datasciencecoursera/ReproducibleResearch/activity.csv")
FullActivityDF<-ActivityDF
ActivityDF<-subset(ActivityDF,!is.na(steps))
DailySteps<-ddply(ActivityDF,.(date), summarize, sum=sum(steps))
```
## Histogram of total number of step taken per day
```{r histogram, echo=TRUE, fig.height=4}
hist(DailySteps$sum,main="Daily Steps Histogram",xlab="Daily Steps",breaks=10)
MeanStepsPerDay<-mean(ActivityDF$steps)
MedianStepsPerDay<-median(ActivityDF$steps)
TotalSteps<-sum(ActivityDF$steps)
````
## What is mean total number of steps taken per day? 
```{r meanmedianperday, echo=TRUE}
ddply(ActivityDF,.(date), summarize, mean=mean(steps),median=median(steps))[,c("date","mean","median")]
```

### Overall Mean Value - `r MeanStepsPerDay`
### Overall Median Value - `r MedianStepsPerDay`

## What is the average daily activity pattern?

```{r averagedailyactivity, echo=TRUE, fig.height=4}
AverageStepsInterval<-ddply(ActivityDF,.(interval), summarize, mean_steps=mean(steps))
plot(AverageStepsInterval$interval,AverageStepsInterval$mean_steps,type="l",
     main="Average Steps Interval Wise", xlab="5 Minutes Interval",ylab='Average Steps Taken')
```


```{r intervalmaxsteps, echo=TRUE, fig.height=4}
IntervalMax<-subset(AverageStepsInterval,mean_steps==max((AverageStepsInterval$mean_steps)))$interval
```
## Invterval with maximum average number of steps across all days - `r IntervalMax`

## Imputing missing values
```{r imputingmissingvalues, echo=TRUE,fig.height=4}
NAActivityDF<-subset(FullActivityDF,is.na(steps))
#Create new data set with replacing NA values with mean steps for the interval
NAActivityDFValues<-merge(NAActivityDF,AverageStepsInterval)[,c("mean_steps","date","interval")]
colnames(NAActivityDFValues)<-c("steps","date","interval")
ActivityDFNew<-rbind(ActivityDF,NAActivityDFValues)
# creating historgram of new activity data
hist(DailySteps$sum,main="Daily Steps Histogram with NA values Replaced",xlab="Daily Steps",breaks=10)
```

```{r meanmedianimputing, echo=TRUE}
MeanStepsPerDayIMV<-mean(ActivityDFNew$steps)
MedianStepsPerDayIMV<-median(ActivityDFNew$steps)
TotalSetpsIMV<-sum(ActivityDFNew$steps)
```
### Mean and Median steps taken per day after imputing missing values with average interval value
### Mean Value - `r MeanStepsPerDayIMV`
### Median Value -`r MedianStepsPerDayIMV`
Mean and Median values are not impacted by imputing missing values
Total steps taken with imputing missing values `r TotalSetpsIMV` is greater than
Total steps taken withoput imputing missing values `r TotalSteps`
## Are there differences in activity patterns between weekdays and weekends?

### More steps are taken during weekend vs weekdays across all intervals

```{r weekdayweekend, echo=TRUE,fig.height=4}
library(lattice)
ActivityDFNew$date1<-as.Date(ActivityDFNew$date,format="%Y-%m-%d")
ActivityDFNew$weekday<-weekdays(ActivityDFNew$date1)
ActivityDFNew$wd_we<-ActivityDFNew$weekday
for (i in 1:nrow(ActivityDFNew)) {
        if (ActivityDFNew[i,]$weekday %in% c("Saturday","Sunday"))
        {
                ActivityDFNew[i,]$wd_we="weekend"
        }
        else 
        {
                ActivityDFNew[i,]$wd_we="weekday" 
        }
}
wd_we_avg_steps<-ddply(ActivityDFNew,.(wd_we,interval), summarize, mean=mean(steps))
wd_we_avg_steps<-transform(wd_we_avg_steps,wd_we=factor(wd_we))
xyplot(mean~interval | wd_we,data=wd_we_avg_steps, layout=c(1,2),ylab="Number of Steps",type="l")
```
